D3064R0: How to Avoid OOTA Without Really Trying

### 2024-02-21 ootavacuous.tex
Alan Stern February 21 feedback, Section 8.1
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

On real systems, instructions require time to execute, even
when executed speculatively.
For example, in Listing~\ref{lst:Speculated Store and Non-Speculated Load},
suppose that the storage for \co{x} is located immediately before that
for \co{y[]}.
Suppose further that hardware speculation incorrectly guesses the value
of \co{i} to be -1, so that the speculated store of 42 on line~9 uses
the address of \co{x}.\footnote{
	Yes, \co{y[-1]} is UB, but the CPU neither knows
	nor cares, nor should it.}
Then line~11 can cause the value 42 to be loaded into \co{r1}, when
by the rules of the C++ abstract machine the value must instead be
either zero or one.

@@

Applying this to Listing~\ref{lst:OOTA Cycle}, because values from
speculative stores cannot be visible to nonspeculative loads, lines~10
and~12 cannot be committed until after the load in line~8 completes
and the conditional on line~9 is evaluated nonspeculatively.\footnote{
	Alternatively, and as noted earlier, one could imagine that both
	the speculation and the squashing of failed speculation might
	span multiple threads.
	This might make sense for multiple hardware threads sharing
	a core.}
Similarly, lines~18 and~20 cannot be committed until after lines~16
and~17 complete nonspeculatively.

>>>>

Computer systems are expected to continue increasing their use of
speculative execution.
Nevertheless, the main points of this paper will remain unaffected.
To see this, keep in mind that on real systems, observable effects
must eventually be committed, but that nothing can be committed
until all of the computations that determine those effects remain
speculative.
Furthermore, instructions take time to execute, even when executed
speculatively.
For example, in Listing~\ref{lst:Speculated Store and Non-Speculated Load},
suppose that the storage for \co{x} is located immediately before that
for \co{y[]}.
Suppose further that hardware speculation incorrectly guesses the value
of \co{i} to be -1, so that the speculated store of 42 on line~9 uses
the address of \co{x}.\footnote{
	Yes, \co{y[-1]} is UB, but the CPU neither knows
	nor cares, nor should it.}
Then line~11 can cause the value 42 to be loaded into \co{r1}, when
by the rules of the C++ abstract machine the value must instead be
either zero or one.

@@

Applying this to Listing~\ref{lst:OOTA Cycle}, because values from
speculative stores cannot be visible to nonspeculative loads, lines~10
and~12 cannot be committed until after the load in line~8 completes
and the conditional on line~9 is evaluated nonspeculatively.
Similarly, lines~18 and~20 cannot be committed until after lines~16
and~17 complete nonspeculatively.
